---
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---
### BAN 502 - Mod 4, Assignment 3
### Meredith, Breandan


```{r Libraries, include=FALSE}
#install.packages("ranger")
#install.packages("gridExtra")
#install.packages("vip")

library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
library(skimr)
```


```{r Loading_the_Data}
drug = read_csv("C:/Users/icema/Desktop/UNCW/BAN 502 (Predictive Analytics)/Module 4 (Other Classification Methods)/Assignment 3/drug_data-1.csv")

names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity", "Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive", "SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis", "Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh", "LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA") 

#str(drug)

drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No" 
drug[drug == "CL2"] = "Yes" 
drug[drug == "CL3"] = "Yes" 
drug[drug == "CL4"] = "Yes" 
drug[drug == "CL5"] = "Yes" 
drug[drug == "CL6"] = "Yes"

drug_clean = drug %>% 
  mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>% 
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44", "45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels = c("Under16", "At16", "At17", "At18", "SomeCollege", "ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>% 
  mutate(Country = factor(Country, labels = c("USA", "NewZealand", "Other", "Australia", "Ireland","Canada","UK"))) %>% 
  mutate(Ethnicity = factor(Ethnicity, labels = c("Black", "Asian", "White", "White/Black", "Other", "White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>% select(-ID)

#str(drug_clean) 

drug_clean = drug_clean %>%
  select(!(Alcohol:Mushrooms)) %>% 
  select(!(Semer:VSA)) 

#names(drug_clean)
```


```{r Task_1}
#skim(drug_clean)
```


*Task 1:* After performing all the necessary cleaning and skimming the data, it appears that there is no missingness within the data. 


```{r Task_2}
set.seed(1234)
drug_clean_split = initial_split(drug_clean, prob = 0.70, strata = Nicotine) 
train = training(drug_clean_split)
test = testing(drug_clean_split)
```


```{r Task 3 p1}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + 
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + 
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + 
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))

grid.arrange(p1,p2,p3,p4, ncol = 2)
```


*Task 3:* Based on the first 4 visualizations, all four variables seem significant in relation to nicotine use. Per the graphs, younger generations are more likely to use nicotine products, specifically those of whom are male. Education level has a notable effect in that the more educated a person is, the less likely they are to use nicotine. The amount of nicotine use also depends on country of origin, with the respondents from the US and Ireland having a higher likelihood of nicotine use than those from other countries such as Canada or the UK. 


```{r Task_3_p5}
p5 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + 
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p6 = ggplot(train, aes(x = Nicotine, y = Nscore)) +
  geom_boxplot()
p7 = ggplot(train, aes(x = Nicotine, y = Escore)) +
  geom_boxplot()
p8 = ggplot(train, aes(x = Nicotine, y = Oscore)) +
  geom_boxplot()

grid.arrange(p5,p6,p7,p8, ncol = 2)
```


*Task 3:* According to the next set of visualizations, ethnicity also seems to be a significant variable in nicotine usage. The respondent's level of extraversion does not seem to have a noticable effect, whereas the respondent's levels of neuroticism and openness to experience do seem to have some sort of effect. 


```{r Task_3_p9}
p9 = ggplot(train, aes(x = Nicotine, y = Ascore)) +
  geom_boxplot()
p10 = ggplot(train, aes(x = Nicotine, y = Cscore)) +
  geom_boxplot()
p11 = ggplot(train, aes(x = Nicotine, y = Impulsive)) +
  geom_boxplot()
p12 = ggplot(train, aes(x = Nicotine, y = SS)) +
  geom_boxplot()

grid.arrange(p9,p10,p11,p12, ncol = 2)
```


*Task 3:* After the final set of visualizations, we can see that the respondent's levels of agreeableness and conscientiousness seem to be slightly varied. There appears to be a much more noticeable effect based on the impulsiveness and sensation-seekingness of the respondent.  


```{r Task_4_folds}
set.seed(123)
rf_folds = vfold_cv(train, v=5)
```


```{r Task_4_random_forest}
drug_clean_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("classification")

drug_clean_wflow =
  workflow() %>%
  add_model(rf_model) %>%
  add_recipe(drug_clean_recipe)

rf_grid = grid_regular(
	mtry(range = c(2,8)),
min_n(range = c(5,20)),
	levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
	drug_clean_wflow,
	resamples = rf_folds,
	grid = rf_grid
)

```


```{r Task_4_visualization}
rf_res_tuned %>%
	collect_metrics() %>%
  filter(.metric == "accuracy") %>%
	mutate(min_n = factor(min_n)) %>% 
	ggplot(aes(mtry, mean, color = min_n)) +
	geom_line(alpha = 0.5, size = 1.5) +
	geom_point() +
  labs(y = "Accuracy")
```


```{r Task_5}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
	drug_clean_wflow,
	best_rf
)

final_rf_fit = fit(final_rf, train)

final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```


*Task 5:* Based on the above information, the most important variables are sensation-seekingness, country (if the respondent is from the UK), the respondent's openness to new experiences, age (45-54), their level of impulsiveness and their conscientiousness. 


```{r Task_6_Training_Set}
trainpredrf = predict(final_rf_fit, train)
#head(trainpredrf)

confusionMatrix(trainpredrf$.pred_class,train$Nicotine, positive = "Yes")
```


*Task 6:* On the training set, the model performs with an accuracy of 88.61%, with a naive accuracy of 67.04%. 


```{r Task_6_Test_Set}
testpredrf = predict(final_rf_fit, test)
#head(testpredrf)

confusionMatrix(testpredrf$.pred_class, test$Nicotine, positive = "Yes") 
```


*Task 6:* On the test set, the model performs with an accuracy of 73.25% and a naive accuracy of 67.09%. Based on the decent-sized drop in accuracy between the training set and testing set, I would say that the model that was made on the training set may be a bit overfit to that specific data and is not as responsive as I would like on new data. 


*Task 7:* In the real world, this model could be used to target pro-recovery and anti-drug-use education and advertisements toward demographics that are more likely to be using certain drugs (in this case, anti-smoking ads toward smokers or those who are likely to try smoking). I would say that this model is usable in the real world, but I would like one that may be a bit more responsive to new data. As the model stands, one of my main concerns is that it may end up targeting too many people who may not be as likely to experiment with drugs. 


```{r}
saveRDS(final_rf_fit, "drug_clean_final_rf_fit.rds")
```

